Output Summary

### 2021-09-02 03-post-process/output-plots.Rmd
Add a table summarizing posterior predictive check results

>>>>

## Posterior Predictive Checks

This output is a diagnostic of the consistency of variability in the data with the variability assumed by the model. For each observed data point and MCMC iteration, a value was simulated from the model. The fit of the observed and simulated data was calculated, and the proportion of MCMC iterations in which the observed data fitted more poorly than the simulated data (i.e., a Bayesian p-value) was calculated. P-values close the 0.5 are ideal (0.4 -- 0.6 have no color below), values closer to 1 (red below) indicate over-dispersion (i.e., model fits real data worse than expected), and values closer to 0 (blue below) indicate under-dispersion (i.e., model fits real data better than expected). Across years within a data set, p-values were summarized as the median and central 95% intervals. The left two columns show the percent of years that had extreme p-values, providing another indication of how often the model fit the real data far better (<0.05) or worse (>0.95) than expected.

The fit criterion differed depending on the likelihood function used to fit the model to the data set:

* **Log-normal**: squared difference of the log-scale observation and log-scale expectation. This applies to all abundance data.
* **Logit-normal**: squared difference of the logit-scale observation and logit-scale expectation. This applies to all survival data, expect pre-spawn survival which used a binomial distribution.
* **Binomial/Multinomial**: chi-squared fit statistic. This applies to all composition data and the pre-spawn survival data.

```{r calculate-bayesian-pvals}
# function to summarize bayesian p-values within a data source
summarize_bp = function(bp) {
  data.frame(
    "N" = apply(bp, 2, function(x) sum(!is.na(x))),
    "median" = apply(bp, 2, median, na.rm = TRUE),
    "q2.5" = apply(bp, 2, function(x) quantile(x, 0.025, na.rm = TRUE)),
    "q97.5" = apply(bp, 2, function(x) quantile(x, 0.975, na.rm = TRUE)),
    "p_lt_0.05" = apply(bp, 2, function(x) sum(x < 0.05 & !is.na(x))/sum(!is.na(x))),
    "p_gt_0.95" = apply(bp, 2, function(x) sum(x > 0.95 & !is.na(x))/sum(!is.na(x)))
  )
}

### CALCULATE BAYESIAN P-VALUES: ABUNDANCE DATA SOURCES ###

# fall trap
new = post_subset(post, "Pa_obs_new_dev", TRUE)
obs = post_subset(post, "Pa_obs_dev", TRUE)
bp = array_format(colMeans(obs > new))
dimnames(bp) = list(all_yrs[1:dim(bp)[1]], "fall-mig", pops)
bp = bp[,"fall-mig",]
Pa_bp = data.frame(
  name = rep("Juveniles", 4), pop = pops,
  origin = rep("NOR", 4), LH_type = rep("Fall Migrants", 4),
  summarize_bp(bp)
)
rownames(Pa_bp) = NULL

# spring trap
new = post_subset(post, "Mb_obs_new_dev", TRUE)
obs = post_subset(post, "Mb_obs_dev", TRUE)
bp = array_format(colMeans(obs > new))
dimnames(bp) = list(all_yrs[1:dim(bp)[1]], c("fall-mig", "spring-mig"), "NOR", pops)
bp = bp[,"spring-mig","NOR",]
Mb_bp = data.frame(
  name = rep("Juveniles", 4), pop = pops,
  origin = rep("NOR", 4), LH_type = rep("Spring Migrants", 4),
  summarize_bp(bp)
)
rownames(Mb_bp) = NULL

# adult abundance
new = post_subset(post, "Ra_obs_new_dev", TRUE)
obs = post_subset(post, "Ra_obs_dev", TRUE)
bp = array_format(colMeans(obs > new))
dimnames(bp) = list(all_yrs[1:dim(bp)[1]], pops)
Ra_bp = data.frame(
  name = rep("Adults", 4), pop = pops,
  origin = rep("Combined", 4), LH_type = rep("Combined", 4),
  summarize_bp(bp)
)
rownames(Ra_bp) = NULL

### CALCULATE BAYESIAN P-VALUES: SURVIVAL DATA SOURCES ###

# summer tagging survival
new = post_subset(post, "Lphi_obs_new_Pb_Ma_dev", TRUE)
obs = post_subset(post, "Lphi_obs_Pb_Ma_dev", TRUE)
bp = array_format(colMeans(obs > new))
dimnames(bp) = list(all_yrs[1:dim(bp)[1]], pops)
phi_Pb_Ma_bp = data.frame(
  name = rep("Summer Tagging to LGR", 4), pop = pops,
  origin = rep("NOR", 4), LH_type = rep("Combined", 4),
  summarize_bp(bp)
)
rownames(phi_Pb_Ma_bp) = NULL

# fall trap survival
new = post_subset(post, "Lphi_obs_new_Pa_Ma_dev", TRUE)
obs = post_subset(post, "Lphi_obs_Pa_Ma_dev", TRUE)
bp = array_format(colMeans(obs > new))
dimnames(bp) = list(all_yrs[1:dim(bp)[1]], "fall-mig", pops)
bp = bp[,"fall-mig",]
phi_Pa_Ma_bp = data.frame(
  name = rep("Fall Trap to LGR", 4), pop = pops,
  origin = rep("NOR", 4),   LH_type = rep("Fall Migrants", 4),
  summarize_bp(bp)
)
rownames(phi_Pa_Ma_bp) = NULL

# trib to lgr survival
new = post_subset(post, "Lphi_obs_new_Mb_Ma_dev", TRUE)
obs = post_subset(post, "Lphi_obs_Mb_Ma_dev", TRUE)
bp = array_format(colMeans(obs > new))
dimnames(bp) = list(all_yrs[1:dim(bp)[1]], c("fall-mig", "spring-mig"), c("NOR", "HOR"), pops)
bp = bp[,"spring-mig",,]
bp = cbind(CAT_NOR = bp[,"NOR","CAT"], CAT_HOR = bp[,"HOR","CAT"],
           LOS_NOR = bp[,"NOR","LOS"], LOS_HOR = bp[,"HOR","LOS"],
           MIN_NOR = bp[,"NOR","MIN"], MIN_HOR = bp[,"HOR","MIN"],
           UGR_NOR = bp[,"NOR","UGR"], UGR_HOR = bp[,"HOR","UGR"])
phi_Mb_Ma_bp = data.frame(
  name = rep("Spring Trap/Release to LGR", 8), pop = rep(pops, each = 2),
  origin = rep(c("NOR", "HOR"), 4), LH_type = rep("Spring Migrants", 8),
  summarize_bp(bp)
)
rownames(phi_Mb_Ma_bp) = NULL

# hydropower survival
new = post_subset(post, "Lphi_obs_new_Ma_O0_dev", TRUE)
obs = post_subset(post, "Lphi_obs_Ma_O0_dev", TRUE)
bp = array_format(colMeans(obs > new))
dimnames(bp) = list(all_yrs[1:dim(bp)[1]], c("NOR", "HOR"))
phi_Ma_O0_bp = data.frame(
  name = rep("Hydrosystem (Downstream)", 2), pop = rep("Combined", 2),
  origin = c("NOR", "HOR"), LH_type = rep("Combined", 2),
  summarize_bp(bp)
)
rownames(phi_Ma_O0_bp) = NULL

### CALCULATE BAYESIAN P-VALUES: COMPOSITION DATA SOURCES ###

# weir composition
new = post_subset(post, "weir_x_obs_new_dev", TRUE)
obs = post_subset(post, "weir_x_obs_dev", TRUE)
bp = array_format(colMeans(obs > new))
dimnames(bp) = list(all_yrs[1:dim(bp)[1]], pops)
bp[jags_data$weir_nx_obs == 0 | is.na(jags_data$weir_nx_obs)] = NA
weir_x_bp = data.frame(
  name = rep("Weir", 4), pop = pops,
  origin = rep("Combined", 4), LH_type = rep("Combined", 4),
  summarize_bp(bp)
)
rownames(weir_x_bp) = NULL

# carcass composition
new = post_subset(post, "carc_x_obs_new_dev", TRUE)
obs = post_subset(post, "carc_x_obs_dev", TRUE)
bp = array_format(colMeans(obs > new))
dimnames(bp) = list(all_yrs[1:dim(bp)[1]], pops)
bp[jags_data$carc_nx_obs == 0 | is.na(jags_data$carc_nx_obs)] = NA
carc_x_bp = data.frame(
  name = rep("Carcasses", 4), pop = pops,
  origin = rep("Combined", 4), LH_type = rep("Combined", 4),
  summarize_bp(bp)
)
rownames(carc_x_bp) = NULL

# pre-spawn survival
new = post_subset(post, "carcs_spawned_new_dev", TRUE)
obs = post_subset(post, "carcs_spawned_dev", TRUE)
bp = array_format(colMeans(obs > new))
dimnames(bp) = list(all_yrs[1:dim(bp)[1]], pops)
carcs_spawned_bp = data.frame(
  name = rep("Prespawn", 4), pop = pops,
  origin = rep("Combined", 4), LH_type = rep("Combined", 4),
  summarize_bp(bp)
)
rownames(carcs_spawned_bp) = NULL

### COMBINE INTO ONE TABLE ###
bp_tab = rbind(
  Pa_bp, Mb_bp, Ra_bp,
  phi_Pb_Ma_bp, phi_Pa_Ma_bp, phi_Mb_Ma_bp,
  phi_Ma_O0_bp, weir_x_bp, carc_x_bp, carcs_spawned_bp
)

# add a main data category variable
category = rep(NA, nrow(bp_tab))
category = ifelse(bp_tab$name %in% c("Adults", "Juveniles"), "Abundance", category)
category = ifelse(bp_tab$name %in% c("Weir", "Carcasses"), "Composition", category)
category[is.na(category)] = "Survival"
bp_tab = data.frame(category = category, bp_tab)

# drop the several cases with no data
# MIN HOR survival to LGR and MIN weir data
bp_tab = bp_tab[bp_tab$N > 0,]

# re-order output/misc formatting
bp_tab = bp_tab[order(bp_tab$category, bp_tab$name, bp_tab$pop, bp_tab$origin),]
bp_tab = bp_tab[,c("category", "name", "pop", "LH_type", "origin", "N", "median", "q2.5", "q97.5", "p_lt_0.05", "p_gt_0.95")]
bp_tab$p_lt_0.05 = paste0(round(bp_tab$p_lt_0.05 * 100), "%")
bp_tab$p_gt_0.95 = paste0(round(bp_tab$p_gt_0.95 * 100), "%")
rownames(bp_tab) = NULL
```

```{r bayesian-pvals-table}
# a function to color cells depending on the value
assign_color = function(x) {
  red_pal = colorRampPalette(c("white", "red"))
  blue_pal = colorRampPalette(c("blue", "white"))
  red_cols = red_pal(5)
  blue_cols = blue_pal(5)
  cols = c(blue_cols, red_cols)
  bin_names = levels(cut(rbeta(1e6, 1, 1), breaks = seq(0,1,0.1), include.lowest = TRUE))
  names(cols) = bin_names
  bins = cut(x, breaks = seq(0,1,0.1), include.lowest = TRUE)
  cols_use = cols[as.character(bins)]
  alpha(unname(cols_use), 0.5)
}

# produce the kable
bp_tab %>%
  kbl(digits = 2, col.names = c("Class", "Lifestage", "Population", "Migratory Type", "Origin", "Years w/Data", "Median", "2.5%", "97.5%", "<0.05", ">0.95"),
      align = c("lllllcccccc")) %>%
  kable_styling(full_width = FALSE, bootstrap_options = c("condensed")) %>%
  add_header_above(c("Data Type Information" = 5, "P-value Summaries" = 4, "% of Extreme P-values" = 2)) %>%
  column_spec(7, background = assign_color(bp_tab[,7])) %>%
  column_spec(8, background = assign_color(bp_tab[,8])) %>%
  column_spec(9, background = assign_color(bp_tab[,9])) %>%
  collapse_rows(1:5) 
```
